---
title: "Homework 6"
author: "Shaolei Ma"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(latex2exp)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

Use the data cleaning procedure similar to HW5. Omit certain cities as instructed.

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = factor(
      case_when(
        disposition == "Closed without arrest" ~ "0",
        disposition == "Open/No arrest"        ~ "0",
        disposition == "Closed by arrest"      ~ "1"
    ),
    labels = c("unsolved", "solved"))
  ) |>  
  filter(
    city_state != "Tulsa, AL",
    city_state != "Dallas, TX",
    city_state != "Phoenix, AZ",
    city_state != "Kansas City, MO"
  )
```

The resulting dataframe has `r nrow(homicide_df)` entries, on variables that include the victim name, race, age, and sex; the date the homicide was reported; and the location of the homicide. In cleaning, I created a `city_state` variable that includes both city and state, and a `resolution` variable to indicate whether the case was closed by arrest.

For the city of Baltimore, MD, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors.

```{r}
baltimore_fit = 
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, data = _, family = binomial())

baltimore_fit |> 
  broom::tidy() |> 
  mutate(
    conf_low = estimate - 1.96 * std.error,
    conf_high = estimate + 1.96 * std.error,
    OR = exp(estimate),
    OR_conf_low = exp(conf_low),
    OR_conf_high = exp(conf_high)
  ) |> 
  filter(term == "victim_sexMale") |> 
  select(term, starts_with("OR"))
```

The adjusted odds ratio for solving homicides comparing male victims to female victims has an estimated value of 0.415 and a confidence interval [0.318, 0.542].

Now, run `glm` for each of the cities in the dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.
```{r}
OR_sex_df =
  homicide_df |> 
  nest(df = - city_state) |> 
  mutate(
    models = map(df, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, data = df, family = binomial())),
    results = map(models, broom::tidy)
  ) |> 
  select(city_state, results) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |>
  mutate(
    conf_low = estimate - 1.96 * std.error,
    conf_high = estimate + 1.96 * std.error,
    OR = exp(estimate),
    OR_conf_low = exp(conf_low),
    OR_conf_high = exp(conf_high)
  ) |> 
  select(city_state, term, starts_with("OR")) |> 
  arrange(OR)

OR_sex_df |> 
  knitr::kable()
```

Create a plot that shows the estimated ORs and CIs for each city. 
```{r}
OR_sex_df |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = OR, y = city_state)) +
  geom_point() +
  geom_errorbar(aes(xmin = OR_conf_low, xmax = OR_conf_high)) +
  labs(title = "OR Among Cities")
```

It could be concluded that New York, NY has the lowest OR, while Albuquerque, NM has the highest OR in terms of solving homicides comparing male victims to female victims.

# Problem 2
